---
title: "IE6600_Sec05_Group_16_Hackathon"
author: "Akilesh Vishnu Mohan Raj & Ravi Patel"
date: "2022-10-16"
output: html_document
---








# Contents :


## 1  Problem Statement 

## 2  About the Dataset

## 3  Output
        3.1 Total Number of Restaurants in San Francisco Analysis
        3.2 Risk Categories Analysis 
        3.3 Geographical Mapping

## 4  Conclusion 

## 5  References


-----------------------------------------------------------------------------

# 1 :   Problem Statement
 
The San Francisco Department of Public Health (SFDPH) is composed of various 
subdivisions that work together to serve the city of San Francisco. SFDPH 
primary aim is to be protecting, prioritizing, and promoting the health of all 
the citizens in San Francisco.

A report and scoring system for inspections have been created by the Health 
Department.  The Health Inspector assigns a score to the facility based on the 
violations discovered during the inspection. Examples of violations include 
group with a high risk: records specific violations related to the contamination
of surfaces where food encounters it, the adulteration of food products, and the
spread of food-borne illnesses. records specific violations that pose a moderat
e threat to the health and safety of the general public. low risk category: 
records violations that pose no immediate danger to the public's health or 
safety or have a low risk. This dataset contains the score card that the 
inspector will issue, which is kept at the food establishment and is accessible
to the general public. 

We find this information of great value for making us understanding and 
underlying issues and factors behind the inspection and their risk categories 
the SFDPH gets. It Is known that the department of public health have a lot of 
inspections regarding the complaints, change in ownership and much more in
restaurants that is why understanding how the inspections are making the 
restaurant in future, are they having more inspections in the future also.

Restaurant Scores - LIVES Standard
The Health Department has developed an inspection report and scoring system. 
After conducting an inspection of the facility, the Health Inspector calculates 
a score based on the violations observed. Violations can fall into:
high risk category: records specific violations that directly relate to the 
transmission of food borne illnesses, the adulteration of food products and the
contamination of food-contact surfaces.moderate risk category: records specific
violations that are of a moderate risk to the public health and safety.
low risk category: records violations that are low risk or have no immediate
risk to the public health and safety.The score card that will be issued by the 
inspector is maintained at the food establishment and is available to the
public in this dataset.

San Francisco's LIVES restaurant inspection data leverages the LIVES Flattened 
Schema 
(https://goo.gl/c3nNvr)
which is based on LIVES version 2.0, cited on 
Yelp's website 
(http://www.yelp.com/healthscores).


--------------------------------------------------------------


# 2 :   About the Dataset


Restaurant Scores dataset contains all the restaurants, their name, their 
location, if they had inspections, their risk categories after inspection and 
much more. The information here is collected by the San Francisco Department
of Public Health. 

Some of the variables in the dataset are as follows:

•	Business name, id, and location  
•	Inspection date, id, and type
•	Violation id, and description
•	Risk category...etc

Source:
https://data.sfgov.org/Health-and-Social-Services/Restaurant-Scores-LIVES-Standard/pyih-      qa8i/data(Found via SFO Open Data)




```{r,echo=FALSE, warning=FALSE,include=FALSE}


library(tidyverse) # metapackage with lots of helpful functions
library(lubridate)
library(janitor)
library(tidyr)
library(dplyr)
library(ggplot2)
library(plotly)
library(readr)
library(ggmap)
library(treemap)
library(treemapify)
library(magrittr)
library(gridExtra)
library(scales)




```




--------------------------------------------------------------



```{r,echo=FALSE, warning=FALSE}

df<-read.csv("Restaurant_Scores.csv",header=TRUE,sep = ",",na.strings = c(""))



```


# 3 : Output


# Section 3.1 Total Number of Restaurents in San Francisco:-


Analysing the DataSet with the Total Number of Restaurents in San Francisco 
and the number of Violations over the Years



# Plot 1



Firstly Let us Observe the total number of business that existed in the State of
California, more in accordance with the city San Francisco as available in the 
San Francisco Department of Public Health (SFDPH) is composed of various 
subdivisions that work together to serve the city of San Francisco.

Thus the below Bar plot says the total number of Restaurants that is present in
the city of San Francisco during the years 2016, 2017, 2018, 2019. Now this is
a very important data point to analyse as only after getting this data we will
be able to do further analysis on the no of restaurants that have violated the
standards and the risk categories that it has been awarded for the respective 
years.


```{r,echo=FALSE, warning=FALSE}

df_1<- df %>% 
       mutate(inspection_date =mdy_hm(inspection_date),
            Month= month(inspection_date), Year=     year(inspection_date)) %>% 
  select(business_id,risk_category,inspection_score,Year) %>% 
  drop_na() %>% 
  group_by(Year) %>% 
  count(Year,name = "Total_Businesses_existed_over_years")

#----------------------------------------------------

ggplot(df_1, aes(x=Year, y=Total_Businesses_existed_over_years, fill=Total_Businesses_existed_over_years))+ geom_bar (stat="identity")+
  geom_text(aes(label=Total_Businesses_existed_over_years), vjust=1.6,hjust=1.5, color="white")+
              coord_flip()



```






```{r,echo=FALSE, warning=FALSE}

df_2<- df %>% 
       mutate(inspection_date =mdy_hm(inspection_date),Month= month(inspection_date), Year=     year(inspection_date)) %>% 
  select(business_id,Year) %>% 
  drop_na() %>% 
  #pivot_wider(names_from =  risk_category,values_from = inspection_score)
  group_by(Year) %>% 
  count(Year) %>% 
  rename("Total_Restaurents" = n ) 
  
#-----------------------------------------

plot_2.1 <- ggplot(df_2, aes(area=Total_Restaurents, label=Total_Restaurents, fill=Total_Restaurents)) +
           geom_treemap() +
           ggtitle("Total Restaurants in SF") +
           geom_treemap_text(fontface = "italic", colour = "white", place = "centre", grow = FALSE)+
  theme_bw()
            
#-----------------------------------------

plot_2.2<-ggplot(df_2, aes(x = "", y = Total_Restaurents, fill = Total_Restaurents)) +
  geom_col(color="white") +
  geom_text(aes(label = Year),
            position = position_stack(vjust = 0.5))+
  coord_polar(theta = "y")+
  theme_bw()







```



```{r,echo=FALSE, warning=FALSE}
df_3<- df %>% 
       mutate(inspection_date =mdy_hm(inspection_date),Month= month(inspection_date), Year=     year(inspection_date)) %>% 
  select(Year,violation_id) %>% 
  drop_na() %>% 
  #pivot_wider(names_from =  risk_category,values_from = inspection_score)
  group_by(Year) %>% 
  count(Year) %>% 
  rename("Violated_Restaurents" = n )

#-----------------------------------------

plot_3.1 <- ggplot(df_3, aes(area=Violated_Restaurents, label=Violated_Restaurents, fill=Violated_Restaurents)) +
           geom_treemap() +
           ggtitle("Restaurents that have Violation codes in SF") +
           geom_treemap_text(fontface = "italic", colour = "white", place = "centre", grow = FALSE)+
  theme_bw()
            
#-----------------------------------------

plot_3.2<-ggplot(df_3, aes(x = "", y = Violated_Restaurents, fill = Violated_Restaurents)) +
  geom_col(color="white") +
  geom_text(aes(label = Year),
            position = position_stack(vjust = 0.5))+
  coord_polar(theta = "y")+
  theme_bw()

```




# Plot_2 :



Now it becomes very crucial to understand and analyse the ratio of total 
restaurants to the no of restaurants that have violated the standards. So the 
below Tree map (1) On the top shows the total number of restaurants across the 
years (2016-2019) and the total violations categorized from (2016-2019)




```{r,echo=FALSE, warning=FALSE}
grid.arrange(plot_2.1, plot_3.1,ncol = 1.5)

```



# Plot 3



The below Pie chart gived more context to the one observed earlier through the 
Tree map by giving the Total Restaurants in San-Francisco Vs Total Violated 
Restaurants in San Francisco from the years (2016-2019).



```{r,echo=FALSE, warning=FALSE}
grid.arrange(plot_2.2, plot_3.2,ncol = 1.5)

```





# Plot 4



Interesting Analogy by Taking the percentage of the total Restaurants vs 
Violated restaurants to find some trend and we can figure out that every 
successive year it is getting low and the next year it is raising and we can 
say that the restaurants becomes reluctant and do not maintain proper measures 
once after the first year and after further scrutinize the next year they
improve in quality



```{r,echo=FALSE,warning=FALSE}


df_4 <- inner_join(df_2,df_3, by = "Year")%>% 
  mutate(Percentage_Violated=(Violated_Restaurents/Total_Restaurents)*100)


plot_4 <- ggplot(df_4, aes(x = Percentage_Violated, fill=Percentage_Violated)) +
           geom_area(stat="bin",position="fill") +
           ggtitle("Stacked Area Graph for Comparison + Part to Whole")

df_4




```




# Some important Questions that have arised during the analysis and the answers for the are:



A) Have the Restaurant Count increased over the Years?

ans) Yes, The restaurants count have drastically inclined over the years.

B) Have the inspections reduced due to the new restaurants coming in?

ans) No, from the analysis it is clear that despite the count of new restaurants 
the inspections has stayed the same and the Government is doing well to keep 
up with the quality standards.




--------------------------------------------------------------



# Section 3.2:- Risk Categories Analysis :

Deep Analysis on the risk Categories of violations over years

This section is a very crucial part of the analytics as it is the deciding 
factor if the restaurant is a good restaurant or a bad restaurant . This analysis
is done based on a lot of factors like below.

Violations can fall into:

High risk category: records specific violations that directly relate to the 
transmission of food borne illnesses, the adulteration of food products and the 
contamination of food-contact surfaces.

Moderate risk category: records specific 
violations that are of a moderate risk to the public health and safety.

Low risk category: records violations that are low risk or have no immediate 
risk to the public health and safety.




# Plot 5

The below Graph shows us the Average Violation scores of the various risk 
categories over the years (2016-2019).

Average Score is the score that lands with a geom_point.

The Risk categories is categorized on the legend.









```{r,echo=FALSE, warning=FALSE}

df_5<- df %>% 
       mutate(inspection_date =mdy_hm(inspection_date),Month= month(inspection_date), Year=     year(inspection_date)) %>% 
  select(inspection_score,risk_category,Year) %>% 
  drop_na() %>% 
  group_by(risk_category,Year) %>% 
  summarise(risk_category,Average_Scores_of_Years=mean(inspection_score),Year) %>% 
  unique()

#-----------------------------------------

plot_5 <- ggplot(df_5, aes(x=Year, y=Average_Scores_of_Years,group=Year,color=risk_category)) +
           geom_line() +
            geom_point(size=2)+
           ggtitle("Average inspection scores of the different Risk Categories")

plot_5


```



# Plot 6



The Below Plot helps us understand how many violations have been categorized 
for the restaurents in the city of San Francisco based on the 
Risk Category Vs Years.

So based on the understanding on the below Groups Bar plot we can interpret 
how many High risk to Low risk restaurants were classified and which year 
the performance was bad and do deep analysis on that particular year.





```{r,echo=FALSE, warning=FALSE}


df_6<- df %>% 
       mutate(inspection_date =mdy_hm(inspection_date),Month= month(inspection_date), Year=     year(inspection_date)) %>% 
  select(risk_category,inspection_score,Year) %>% 
  drop_na() %>% 
  group_by(risk_category,Year) %>% 
  count(risk_category,Year,name = "Categorized_Violation_Count") 
  #pivot_wider(names_from = Year,values_from = Inspection_Count)

#-----------------------------------------

ggplot(data=df_6, aes(x=Year, y=Categorized_Violation_Count, fill=risk_category)) +
      geom_bar(stat="identity", position=position_dodge())+
      scale_fill_brewer(palette="Blues") +
      geom_text(aes(label=Categorized_Violation_Count), vjust=1.6, color="white",
      position = position_dodge(0.8), size=3.5)

```





# Plot 7



Average Categorical Inspection Scores over the years which is the expected 
metrix for quantifying the volume of violation over the various Risk Categories.

We have acheived in computing this with a line chart and an Area Plot.




```{r,echo=FALSE, warning=FALSE,dplyr.summarise.inform = FALSE}

df_7<- df %>% 
       mutate(inspection_date =mdy_hm(inspection_date),Month= month(inspection_date), Year=     year(inspection_date)) %>% 
  select(risk_category,inspection_score) %>% 
  drop_na() %>% 
  group_by(risk_category) %>% 
  summarise(Category_Average=mean(inspection_score)) %>% 
  arrange(-Category_Average)

#-----------------------------------------
  
df_7.1<- df %>% 
       mutate(inspection_date =mdy_hm(inspection_date),Month= month(inspection_date), Year=     year(inspection_date)) %>% 
  select(risk_category,inspection_score) %>% 
  drop_na() %>% 
  group_by(risk_category)

#-----------------------------------------


plot_7 <- ggplot(df_7, aes(x=risk_category, y=Category_Average,group=1)) +
           geom_line(color="blue") +
            geom_point()+
           ggtitle("Average inspection scores of the different Risk Categories")

#-----------------------------------------
plot_7.1 <- ggplot(df_7.1, aes(x = inspection_score, fill=risk_category)) +
           geom_area(stat="bin",position="fill") +
           ggtitle("Inspection Scores over the Risk Categories")


#-----------------------------------------

grid.arrange(plot_7, plot_7.1,ncol = 1.5)

```





# Some important Questions that have arised during the analysis and the answers for the are:



A) Have The High Risk Categories increased/Decreased over years?

ans) There is no Drastic change in the High risk category in a progressive mannar.
But one interesting observation is that every Successive year there is a Dip
and Increase in the next successive year. From this we can Infer that the 
Restaurants are doing good once inspection is done but the next year they are 
missing some quality standards which is resulting in the Hike in High risk ratio.




--------------------------------------------------------------




# Section 3.3: Geographical Mapping of the Violations wrt Risk_Category








```{r,echo=FALSE, warning=FALSE}

# Reformat the date as year into a new variable

rest_score_data <- read.csv("Restaurant_Scores.csv",header=TRUE,sep = ",",na.strings = c(""))

dat_with_year <- rest_score_data %>% 
    mutate(inspection_date =mdy_hm(inspection_date),Month= month(inspection_date), inspection_year=     year(inspection_date))


# Select the variables to be use for plotting

risk_per_year <- select(dat_with_year, inspection_year, risk_category) %>%
    na.omit(risk_category)




```



# Plot 8: 



All Years Risk wise Map:

San Francisco Geographical map of total number of Violations during the years
of (2016-2019). This involves all the risk Categories from low risk, moderate 
risk and High risk hilighted as scatters in the Map with the colours green , 
blue and red as hilighted in the Legend. These colours represent the severity 
of the violation.


```{r,echo=FALSE, warning=FALSE,dplyr.summarise.inform = FALSE}


sf_coord <- c(left = -122.519320, bottom = 37.68, right = -122.35, top = 37.82)
sf <- get_stamenmap(sf_coord, zoom = 13, maptype = "watercolor")

clean_dat_with_year <- select(dat_with_year, business_name, business_longitude, business_latitude, inspection_year, risk_category) %>% na.omit



p2 <- ggmap(sf) +
    geom_point(data = clean_dat_with_year, aes(x = business_longitude, y = business_latitude,text = business_name,col = risk_category)) +
        labs(x = "longitude", y = "latitude") +
        facet_wrap( ~ inspection_year, ncol=2)

p2

```




# Plot 9

All Years only HIGH Risk wise Map:

The below Geographical map of San Francisco gives our more clear emphasis on 
the High risk categories of restaurants over years (2016-2019). Now this is 
very important for the analysis as we have a perfect idea on the location of 
the High risk restaurents.


```{r, echo = F, message = F, warning = F}
high_risk_score <- filter(clean_dat_with_year, risk_category == "High Risk")

p3 <- ggmap(sf) +
    geom_point(data = high_risk_score, aes(x = business_longitude, y = business_latitude,text = business_name,col = "risk_category")) +
        labs(x = "longitude", y = "latitude") +
        facet_wrap( ~ inspection_year, ncol=2)

p3


```




# Some important Questions that have arised during the analysis and the answers for the are:



A) Are the Restaurants equally spread across San Francisco:

ans) No, The restaurants count is more on Towards the Downtown City areas 
and keeps continouesly increasing in those areas. 



-------------------------------------------------------------------------



# Conclusion


Over this research, we were able to learn a lot of things how the Department of 
Public Health in San Francisco works and how are they able to inspect the 
restaurants,reasons for inspection, if there are risky restaurants then they 
provide violation description and categorized them in risk categories.We firmly 
believe that the insights and the visualization provided by data analysis can 
help the restaurants and customers about the quality of a restaurant and 
give improvement measures and thus improve quality of Food,Service and customer
Satisfaction

The first thing that we discovered from the data was that major of the 
restaurant had a lot of inspections may be in different branches and each year 
the number of business keep on increasing on exponential scale and a lot of 
restaurant had opened up new branches in different locations.


Section - 3.1:

The First plot tells us that the total number of business in San Francisco keeps
on increasing every year.The data is shown from 2016-2019 years.This was the 
basis for us to further dive into data and analysis it thoroughly.

The Second plot and Third plot  tells us the comparison between the total number
of restaurants and the number of restaurants that have violated using pictorial
representation in the form of tree map and pie chart over the years.

Section - 3.2:

The Fourth plot tells us the comparison between the percentage of restaurants 
and the violated restaurants. One can see here that if the restaurant has 
violated in the current year then it will improve its service quality in the 
coming year

The Fifth plot tells us the average inspection score of the restaurant in each 
year the data has been recorded.

The Sixth plot tells us the categorized violation count for each risk category 
in each year.

The Seventh plot tells us the difference between the average inspection scores 
for different risk categories and the stacked are graph tell us the comparison
between different risk categories and the inspection score


Section - 3.3:

The Eight plot tell us the total number of violations in each year with each 
risk categories that is scattered across the geographical representation of 
San Francisco. Also we have two Geographical Graphs one that sums all the available 
restaurants across all risk categories and the next gives more context on the 
restaurants with High Risk.



The Department of Public Health of San Francisco were trying their best to 
inspect the restaurant that were violating guidelines and taking strict actions 
as much as possible. We also learned that the department was listening and 
coorporating with the citizins if they had issues regarding the restaurants.
We can Infer that every Successive year there is a Dip
and Increase in the next successive year. From this we can Infer that the 
Restaurants are doing good once inspection is done but the next year they are 
missing some quality standards which is resulting in the Hike in High risk ratio.




------------------------------------------------------------------------

# References

DataSet: 
https://data.sfgov.org/Health-and-Social-Services/Restaurant-Scores-LIVES-Standard/pyih-      qa8i/data(Found via SFO Open Data)




























